Stack Safety as a Security Property

### 2020-07-05 main.tex
Maybe call things MO and MPO

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  Abstract formal properties capturing the
  intuitive concepts of local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in a very general setting. These properties
  rule out common examples of stack memory abuse.
\item
  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
  \apt{If we have no testing, do we even want to talk about the testing
    property?} \bcp{Not here; just in future work.  I guess we need to see
    whether we wind up with any results at all about testing.}
\item
  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our property, but we give a
  variant of the policy that does enforce a (provably) weaker variant of our
  property.
%
  \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
    but the point we can make is still strong: To make the property easier
    to enforce, we can slightly weaken its guarantees; the weaker version is
    a bit more complicated to state but still retains the pragmatic
    protections of the original.}\apt{To clarify: Nick's LT&C policy uses
    depth-based tags, which do not enforce even our lazy policy. So we need to
    introduce a version of LT&C that uses unique activation tags.}

@@

A trace over a type $A$ is a nonempty, finite or infinite sequence of
$A$s. We will mostly be concerned with sequences of machines states,
denoted $\machT$, and with sequences of triples of (machine states,
policy states, and observations), usually denoted $\MPT$.  \rb{The way
  the definitions are set up, one may expect MPO, but\ldots} We
will require four generic operations on traces.

@@

The ``trace-of'' operator, written \(\hookrightarrow\), coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:
%
\[\mach \hookrightarrow \machOT : \MACHS \rightarrow \MOTRACE\]%
%
\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machT\)}%
%
\noindent
Since the step function is total, if \(\mach \hookrightarrow \machT\),
then \(\machT\) must be infinite.

@@

In a similar fashion, we define \(\hookrightarrow_P\) to relate a
machine-policy state pair to the trace of triples of machine states,
policy states, and observations induced by by \(\stepstoobs{}_P\).
Unlike the $\MOTRACE$ above, this trace can be finite as the
policy can cause a fault:
%
\[(m,p) \hookrightarrow_P MP : \MACHS \times \POLS \rightarrow \MPOTRACE \]
%
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
%
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machT\)}
We will also use \(\pi_m\) to project the trace of machine states
out of a trace, \(\pi_p\) to project the trace of policy states, and
\(\pi_o\) to project the observations.  \bcp{I'm getting a little lost in
  all these.  E.g., do these traces include the final state of a finite
  execution?  If so, what observation is it paired with?}\leo{The final
  observation is a $\tau$ based on the rule in the inlined figure, so
  it is essentially ignored. Point that out somehow?}

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
Let \(f\) be a predicate on machine states. Then we define \(\mach
\hookrightarrow \machT | f\) (read ``\(\machT\) is the prefix of
running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I
  think.}  \leo{There were inconsistencies which i fixed. I wanted the
  metavariables defined to be consistent and the ``temporary''
  variable which is the full trace to be primed. Should I swap those?}

\judgmenttwo{\(\mach \hookrightarrow \machT'\)}
            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machT')
              = \machT\)}
            {\(\mach \hookrightarrow \machT | f\)}
\noindent
And similarly for traces with policies:

@@

      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
      agrees with \(\mach_e\) on every component the callee should not change,
      but with \(\mach_r\) on those components that it is permitted to change.
      So the observable behavior that follows is the idealized behavior to
      which we compare the trace from \(\mach_r\), which must prefix it (to
      account for premature halts due to policy violations.)
      %
      \[\begin{split}
        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machT' \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machT')
      \end{split}\]

@@

      A pair of traces \(\MPT\) and \(\machT\) starting from \(\mach_e\) and
      \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\),
      enjoy observable trace confidentiality with respect to \(\contour\) if
      two conditions hold:

      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
      \pi_o(\machT)\).

      If \(\MPT\) is finite and ends with the state
      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machT\) is finite. Let
      \(\mach_r'\) be the final state of \(\machT\). Then the induced traces
      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
      \hookrightarrow \machT'\) are observationally prefixed,
      \(\pi_o(\MPT') \lesssim \pi_o(\machT')\).

>>>>

In summary, we offer the following contributions:
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  Abstract formal properties capturing the
  intuitive concepts of local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in a very general setting. These properties
  rule out common examples of stack memory abuse.
\item
  A case study of the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:enforcement}).
  This policy enforces our stack safety property, and also a (provably) stronger property that is easier to test.
  \apt{If we have no testing, do we even want to talk about the testing
    property?} \bcp{Not here; just in future work.  I guess we need to see
    whether we wind up with any results at all about testing.}
\item
  A refinement of the case study that considers the more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}~(\cref{sec:lazy}).  This policy does not enforce our property, but we give a
  variant of the policy that does enforce a (provably) weaker variant of our
  property.
%
  \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
    but the point we can make is still strong: To make the property easier
    to enforce, we can slightly weaken its guarantees; the weaker version is
    a bit more complicated to state but still retains the pragmatic
    protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
    depth-based tags, which do not enforce even our lazy policy. So we need to
    introduce a version of LT\&C that uses unique activation tags.}

@@

A trace over a type $A$ is a nonempty, finite or infinite sequence of
$A$s. We will mostly be concerned with sequences of pairs of machine states
and observations, called machine traces and denoted $\machOT \in \MOTRACE$,
and with sequences of triples of a machine state, a policy state, and an
observation, denoted $\MPT \in \MPOTRACE$ and referred to as MP-traces. 
\rb{The way the definitions are set up, one may expect MPO, but\ldots}

@@

The ``trace-of'' operator, written \(\hookrightarrow\), coinductively
relates a machine state with the trace of machine states paired with
observations produced by repeated application of step:
%
\[\mach \hookrightarrow \machOT : \MACHS \rightarrow \MOTRACE\]%
%
\judgmenttwo{\(\mach_0 \stepstoobs{\obs} \mach_1\)}{\(\mach_1 \hookrightarrow \machOT\)}
            {\(\mach_0 \hookrightarrow (\mach_0,\obs) \machOT\)}%
%
\noindent
Since the step function is total, if \(\mach \hookrightarrow \machOT\),
then \(\machOT\) must be infinite.

@@

In a similar fashion, we define \(\mpstatename \hookrightarrow_P \MPT\)
to relate a machine-policy state pair to the MP-trace
induced by by \(\stepstoobs{}_P\). Unlike the $\MOTRACE$ above, this trace can
be finite as the policy can cause a fault, in which case the final state is
a \(\tau\):
%
\judgment{\(\neg \exists \pol_1 . \pol_0 \rightharpoonup \pol_1\)}
         {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow \mpostate{\mach_0}{\pol_0}{\tau}\)}
%
\judgmentthree{\(\mach_0 \stepstoobs{\obs} \mach_1\)}
              {\(\pol_0 \rightharpoonup \pol_1\)}
              {\(\mpstate{\mach_1}{\pol_1} \hookrightarrow \machOT\)}
              {\(\mpstate{\mach_0}{\pol_0} \hookrightarrow_P \mpostate{\mach_0}{\pol_0}{\obs} \machOT\)}
We will also use \(\pi_m\) to project the trace of machine states
out of a trace, \(\pi_p\) to project the trace of policy states, and
\(\pi_o\) to project the observations, with \(\pi_m\) and \(\pi_o\) overloaded
for machine traces and MP-traces.

\bcp{I'm getting a little lost in
  all these.  E.g., do these traces include the final state of a finite
  execution?  If so, what observation is it paired with?}\leo{The final
  observation is a $\tau$ based on the rule in the inlined figure, so
  it is essentially ignored. Point that out somehow?}\sna{It is now pointed
  out, but it also reminds me as we go back to prefix-up-to that the final
  observation of a prefix needs to be tau...}

@@

We will frequently take a prefix of the trace from an initial state up
through the first state where some condition holds on the machine state.
This can easily be implemented using \(\PUT\), and we provide special notation
for convenience.
%
Let \(f\) be a predicate on machine states. Then we define \(\mach
\hookrightarrow \machOT | f\) (read ``\(\machOT\) is the prefix of
running \(\mach\) up to \(f\)''): \bcp{M and M' should be swapped, I
  think.}  \leo{There were inconsistencies which i fixed. I wanted the
  metavariables defined to be consistent and the ``temporary''
  variable which is the full trace to be primed. Should I swap those?}

\judgmenttwo{\(\mach \hookrightarrow \machOT'\)}
            {\(\PUT(\lambda (\mach',\obs).f\ \mach',\machOT')
              = \machOT\)}
            {\(\mach \hookrightarrow \machOT | f\)}
\noindent
And similarly for traces with policies:

@@

      If \(\mach_e\) is an entry to a call with contour \(\contour\) and
      \(\mach_r\) its return, \(\mathit{roll}(\contour,\mach_e,\mach_r)\)
      agrees with \(\mach_e\) on every component the callee should not change,
      but with \(\mach_r\) on those components that it is permitted to change.
      So the observable behavior that follows is the idealized behavior to
      which we compare the trace from \(\mach_r\), which must prefix it (to
      account for premature halts due to policy violations.)
      %
      \[\begin{split}
        \text{\sc ObsInt}_\contour ~ \MPT \triangleq
          & \MPT = \mpostate{\mach_e}{\pol_e}{\obs_e} \dots
            \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
        & \mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT' \Rightarrow \\
        & \mathit{roll}(\contour,\mach_e,\mach_r) \hookrightarrow \machOT' \Rightarrow \\
        & \pi_o(\MPT') \lesssim_O \pi_o(\machOT')
      \end{split}\]

@@

      A pair of traces \(\MPT\) and \(\machOT\) starting from \(\mach_e\) and
      \(\mach_e'\), respectively, with \(\mach_e \approx_\contour \mach_r'\),
      enjoy observable trace confidentiality with respect to \(\contour\) if
      two conditions hold:

      The two traces are observationally equivalent: \(\pi_o(\MPT) \eqsim
      \pi_o(\machOT)\).

      If \(\MPT\) is finite and ends with the state
      \(\mpostate{\mach_r}{\pol_r}{\obs_r}\), then \(\machOT\) is finite. Let
      \(\mach_r'\) be the final state of \(\machOT\). Then the induced traces
      \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT'\) and
      \(\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
      \hookrightarrow \machOT'\) are observationally prefixed,
      \(\pi_o(\MPT') \lesssim \pi_o(\machOT')\).
